Data Specialization for a Staged Lambda Calculus

### 2015-02-28 paper/appendix.tex
added color annotations

Our staging annotations indicate that the list of strings are first-stage, while
the query string is second-stage.
\begin{lstlisting} 
atsigngr { 
datatype letter = A | B | C 
datatype string = EmptyS | ConsS of letter * string
datatype list   = EmptyL | ConsL of string * list

fun partition (EmptyL) = (false,EmptyL,EmptyL,EmptyL) 
  | partition (ConsL (s,ss)) = 
  let val (anyEmpty,a,b,c) = partition ss in 
  case s of 
    EmptyS => (true,a,b,c) 
  | ConsS (A,z) => (anyEmpty,ConsL(z,a),b,c) 
  | ConsS (B,z) => (anyEmpty,a,ConsL(z,b),c) 
  | ConsS (C,z) => (anyEmpty,a,b,ConsL(z,c)) 
}
atsignnext { 
datatype letter2 = A2 | B2 | C2
datatype string2 = EmptyS2 | ConsS2 of letter2 * string2
}

fun exists (gr{EmptyL},next{s}) = next{false} 
  | exists (gr{ConsL ht},next{s}) = 
	let atsigngr{
	  val (anyEmpty,a,b,c) = partition (ConsL ht)
	} in next { 
	case s of 
	  EmptyS2 => prev{hold gr{anyEmpty}} 
	| ConsS2 (A2,z) => prev{exists(gr{a},next{z})} 
	| ConsS2 (B2,z) => prev{exists(gr{b},next{z})}
	| ConsS2 (C2,z) => prev{exists(gr{c},next{z})}  
	} 
\end{lstlisting}
Running the splitting algorithm on this code produces:
\begin{lstlisting} 
datatype trie = Leaf 
			  | Branch of bool * trie * trie * trie

1`datatype letter = A | B | C 
datatype string = EmptyS | ConsS of letter * string
datatype list   = EmptyL | ConsL of string * list
fun partition (EmptyL) = (false,EmptyL,EmptyL,EmptyL) 
  | partition (ConsL (s,ss)) = 
  let val (anyEmpty,a,b,c) = partition ss in 
  case s of 
    EmptyS => (true,a,b,c) 
  | ConsS (A,z) => (anyEmpty,ConsL(z,a),b,c) 
  | ConsS (B,z) => (anyEmpty,a,ConsL(z,b),c) 
  | ConsS (C,z) => (anyEmpty,a,b,ConsL(z,c)) 
in
fun exists (EmptyL,()) = ((), Leaf)
  | exists (ConsL ht,()) =
	let 
	  val (anyEmpty,a,b,c) = partition (ConsL ht) 
	in ((), 
		Branch (anyEmpty, 
			#2 (exists (a, ())), 
			#2 (exists (b, ())), 
			#2 (exists (c, ()))))

`2`datatype letter2 = A2 | B2 | C2
datatype string2 = EmptyS2 | ConsS2 of letter2 * string2
fun exists (((),s), Leaf) = false
  | exists (((),s), Branch (e,a,b,c)) = 
    case s of
	  EmptyS2 => e
	| ConsS2 (A,z) => exists (((), z), a)
	| ConsS2 (B,z) => exists (((), z), b)
	| ConsS2 (C,z) => exists (((), z), c)`

>>>>

Our staging annotations indicate that the list of strings are first-stage, while
the query string is second-stage.
\begin{lstlisting} 
3`atsigngr { `1`
datatype letter = A | B | C 
datatype string = EmptyS | ConsS of letter * string
datatype list   = EmptyL | ConsL of string * list

fun partition (EmptyL) = (false,EmptyL,EmptyL,EmptyL) 
  | partition (ConsL (s,ss)) = 
  let val (anyEmpty,a,b,c) = partition ss in 
  case s of 
    EmptyS => (true,a,b,c) 
  | ConsS (A,z) => (anyEmpty,ConsL(z,a),b,c) 
  | ConsS (B,z) => (anyEmpty,a,ConsL(z,b),c) 
  | ConsS (C,z) => (anyEmpty,a,b,ConsL(z,c)) 
`3`}`1`
atsignnext { `2`
datatype letter2 = A2 | B2 | C2
datatype string2 = EmptyS2 | ConsS2 of letter2 * string2
`1`}

fun exists (`3`gr{`1`EmptyL`3`}`1`,next{`2`s`1`}) = next{`2`false`1`} 
  | exists (`3`gr{`1`ConsL ht`3`}`1`,next{`2`s`1`}) = 
	let `3`atsigngr{`1`
	  val (anyEmpty,a,b,c) = partition (ConsL ht)
	`3`}`1` in next { `2`
	case s of 
	  EmptyS2 => prev{`1`hold `3`gr{`1`anyEmpty`3`}`2`} 
	| ConsS2 (A2,z) => prev{`1`exists(`3`gr{`1`a`3`}`1`,next{`2`z`1`})`2`} 
	| ConsS2 (B2,z) => prev{`1`exists(`3`gr{`1`b`3`}`1`,next{`2`z`1`})`2`}
	| ConsS2 (C2,z) => prev{`1`exists(`3`gr{`1`c`3`}`1`,next{`2`z`1`})`2`}  
	`1`}` 
\end{lstlisting}
Running the splitting algorithm on this code produces:
\begin{lstlisting} 
datatype trie = Leaf 
			  | Branch of bool * trie * trie * trie

1`datatype letter = A | B | C 
datatype string = EmptyS | ConsS of letter * string
datatype list   = EmptyL | ConsL of string * list
fun partition (EmptyL) = (false,EmptyL,EmptyL,EmptyL) 
  | partition (ConsL (s,ss)) = 
  let val (anyEmpty,a,b,c) = partition ss in 
  case s of 
    EmptyS => (true,a,b,c) 
  | ConsS (A,z) => (anyEmpty,ConsL(z,a),b,c) 
  | ConsS (B,z) => (anyEmpty,a,ConsL(z,b),c) 
  | ConsS (C,z) => (anyEmpty,a,b,ConsL(z,c)) 
in
fun exists1 (EmptyL,()) = ((), Leaf)
  | exists1 (ConsL ht,()) =
	let 
	  val (anyEmpty,a,b,c) = partition (ConsL ht) 
	in ((), 
		Branch (anyEmpty, 
			#2 (exists1 (a, ())), 
			#2 (exists1 (b, ())), 
			#2 (exists1 (c, ()))))

`2`datatype letter2 = A2 | B2 | C2
datatype string2 = EmptyS2 
                 | ConsS2 of letter2 * string2
fun exists2 (((),s), Leaf) = false
  | exists2 (((),s), Branch (e,a,b,c)) = 
    case s of
	  EmptyS2 => e
	| ConsS2 (A,z) => exists2 (((), z), a)
	| ConsS2 (B,z) => exists2 (((), z), b)
	| ConsS2 (C,z) => exists2 (((), z), c)`
